Boutiques: a flexible framework for automated application integration in computing platforms

### 2017-11-03 boutiquespaper.tex
formatting
Various fixes, including issue #2

We present \boutiques, a system to publish, integrate and execute
containerized command-line applications across platforms (see
Figure~\ref{fig:diagram}).  In \boutiques, a command line is described
using a flexible template comprising the inputs it requires and the
outputs it produces. Inputs may be inter-dependent and passed in
configuration files or directly on the command line. Such formal
descriptions, simply referred to as \emph{descriptors}, link to a
container image where the given application and all its dependencies
are installed. \boutiques descriptors allow for automatic application
integration in platforms, and advanced validation of input values to
prevent errors. \boutiques also provides a simple parallelization
framework that allows parallel applications to concurrently execute
multiple tasks (in the remainder, tasks are called
\emph{invocations}). \boutiques descriptors are intented to be
produced by scientific application developers, stored alongside their
application, indexed by common repositories, and consumed by execution
platforms.  A set of core tools facilitate the construction,
validation, import, execution, and publishing of \boutiques
descriptors.
\begin{figure}
  \includegraphics[width=\columnwidth]{boutiquesfig1.pdf}
  \caption{Publication, integration and execution of applications with \boutiques.}
  \label{fig:diagram}
\end{figure}

@@

%\note{Tristan}{Vocabulary (check consistency across the paper): tool vs. application}
%\note{Tristan}{Targeted journals: Scientific Data, Giga Science, Frontiers in Neuroinformatics, Future Generation Computer Systems}

@@

CBRAIN (\url{http://cbrain.mcgill.ca}) ~\cite{SHER-14} is a web
platform that can process data distributed into multiple storage
locations on computing clusters and clouds. CBRAIN offers transparent
access to remote data sources, distributed computing sites, and an
array of processing and visualization tools within a controlled,
secure environment.  The CBRAIN service deployed at the Montreal
Neurological Institute relies on the infrastructure provided by
Compute Canada~\cite{das2016mni}. It currently provides 500+
collaborators in 22 countries with web access to several systems,
including 6 clusters of the Compute Canada high-performance computing
infrastructure (totaling more than 100,000 computing cores and 40PB of
disk storage) and Amazon EC2. CBRAIN transiently stores about 10
million files representing over 50TB distributed in 42 servers. 51
public data processing applications are integrated and over 340,000
processing batches have been submitted since 2010.

@@

We also extended CBRAIN to enable the parallelization of workflows
wrapped as \boutiques descriptors. Applications with the
\texttt{cbrain:can-submit-new-tasks} custom property may submit sub-tasks by
creating \boutiques invocations in their working directory. CBRAIN
periodically scans working directories, submits the requested
sub-tasks, and writes back the invocation id in the same
directory. This parallelization model is simple, the application only
needs to write \boutiques invocations and communication happens
through the file system, and it is also powerful as it enables the
parallelization of complex workflows such as the FSL and Niak ones
described later.

>>>>

We present \boutiques, a system to publish, integrate and execute
containerized command-line applications across platforms (see
Figure~\ref{fig:diagram}).  In \boutiques, a command line is described
using a flexible template comprising the inputs it requires and the
outputs it produces. Inputs may be inter-dependent and passed in
configuration files or directly on the command line. Such formal
descriptions, simply referred to as \emph{descriptors}, link to a
container image where the given application and all its dependencies
are installed. \boutiques descriptors allow for automatic application
integration in platforms, and advanced validation of input values to
prevent errors. \boutiques also provides a simple parallelization
framework that allows parallel applications to concurrently execute
multiple tasks (in the remainder, tasks are called
\emph{invocations}). \boutiques descriptors are intented to be
produced by scientific application developers, stored alongside their
application, indexed by common repositories, and consumed by execution
platforms.  A set of core tools facilitate the construction,
validation, import, execution, and publishing of \boutiques
descriptors.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{boutiquesfig1.pdf}
  \caption{Publication, integration and execution of applications with \boutiques.}
  \label{fig:diagram}
\end{figure}

@@

CBRAIN (\url{http://github.com/aces/cbrain}) ~\cite{SHER-14} is a web
platform that can process data distributed into multiple storage
locations on computing clusters and clouds. CBRAIN offers transparent
access to remote data sources, distributed computing sites, and an
array of processing and visualization tools within a controlled,
secure environment.  The CBRAIN service deployed at the Montreal
Neurological Institute relies on the infrastructure provided by
Compute Canada~\cite{das2016mni}. It currently provides 500+
collaborators in 22 countries with web access to several systems,
including 6 clusters of the Compute Canada high-performance computing
infrastructure (totaling more than 100,000 computing cores and 40PB of
disk storage) and Amazon EC2. CBRAIN transiently stores about 10
million files representing over 50TB distributed in 42 servers. 51
public data processing applications are integrated and over 340,000
processing batches have been submitted since 2010.

@@

We also extended CBRAIN to enable the parallelization of workflows
wrapped as \boutiques descriptors. Applications with the
\texttt{cbrain:can-submit-new-tasks} custom property may submit sub-tasks by
creating \boutiques invocations in their working directory. CBRAIN
periodically scans working directories, submits the requested
sub-tasks, and writes back the invocation id in the same
directory. This parallelization model is simple, the application only
needs to write \boutiques invocations and communication happens
through the file system, and it is also powerful as it enables the
parallelization of complex workflows such as the Niak ones
described later.
